---
title: "01_OSRS_GE_EDA"
author: "Felix Gabay"
format: html
---

# Problem Statement

Players in the popular MMORPG Old School Runescape often flip commodities on the
game's auction house (the Grand Exchange, or 'GE') for profit.

The age-old advice for newcomers to the flipping market is short and simple.

Buy low, sell high. 

Current strategies for flipping on the GE are mostly heuristic, relying on 
experimentation and data collection until a player finds their rhythm and 
preferred commodities. I have personally explored flipping to some degree, yet
never found enough interest (nor profit) to continue my endeavors. I can't help
but wonder, **does there exist a theoretical best portfolio in terms of profit
margin and sale volume?**


## Sierra 117

Let's start by loading up some packages and the dataset I'll be using.

```{r}
# Packages
packages <- ("tidyverse")

for(pkg in packages) {
  if(!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
  library(pkg, character.only = TRUE)
}
```

## Exploratory Data Analysis

### Understand the problem and data

- What is the goal or problem we are trying to solve?
- What variables are present in the dataset and what do they represent?
- What types of data are available?
- Are there any data quality issues or limitations?

### Import and inspect the data

- Load the dataset
- Check numbers of rows and columns
- ID missing values
- Verify data types
- Look for errors, invalid values, or unusual data points

```{r}
# Data
GE_snapshot_2026_06_27 <- read_csv("../data/GE_snapshot_2026-06-27.csv")
head(GE_snapshot_2026_06_27)
```

```{r}
# Start with summary function
summary <- GE_snapshot_2026_06_27 |>
  group_by(item_id, name) |>
  summarize(
    'count' = n(),
    'min_time' = min(timestamp),
    'max_time' = max(timestamp),
    'diff_days' = difftime(max(timestamp), min(timestamp), units = "days"),
    across(
      .cols = all_of(c(
        'avgHighPrice',
        'avgLowPrice',
        'highPriceVolume',
        'lowPriceVolume'
      )),
      .fns = list(mean = mean, sd = sd),
      .names = "{fn}_{col}"
    )
  )

view(summary)
```

### Handle missing data
Why is the data missing?
- MCAR, MAR, MNAR
- [Types of missing data: MCAR, MAR, and MNAR explained](https://learndata.xyz/posts/types%20of%20missing%20data%20mcar,%20mar,%20and%20mnar%20explained/)
- **Understanding this is key to determining the right approach, as it could introduce biases**
- Remove of fill missing values?
- Use suitable imputation methods

### Explore data characteristics
- Check data distribution to understand how values are spread across the dataset
- Measure central tendency using mean, median and mode to find the typical value of the data
- Measure variability using standard deviation to see how much the values vary
- Analyze distribution shape using skewness and kurtosis
- Identify outliers or anomalies that may affect the analysis

### Perform data transformation
Prepare the dataset for better analysis and modeling
- Scaling or normalizing numerical variables
- Encoding categorical variables (one-hot encoding, label encoding, factoring)
- Apply mathematical transformations to correct skewness or non-linearity
- Feature engineering by deriving useful information from existing variables
- Aggregating or grouping data based on specific variables or conditions

### Visualize data relationships
- Bar charts and pie charts for analyzing categorical data distribution
- Histograms, box plots and density plots show distribution and detect outliers in data
- Scatter plots and correlation measures help analyze relationships between variables

### Handle outliers
- Identify extreme values with statistical methods like Interquartile Range or Z-score
- Analyze outliers carefully before taking action
- Use domain knowledge to determine whether they are valid or erroneous
- Apply techniques like capping or transformation if necessary
- Remove outliers only when they are clearly incorrect or harmful to a valid analysis

### Communicate findings and insights
- State the goal and scope of the analysis
- Provide background or context so the approach is easy to understand
- Use visualizations to support findings and make results clearer
- Highlist key insights, patterns, or challenges faced during the analysis
- Suggest next steps or areas that require further investigation






