---
title: "01_OSRS_GE_EDA"
author: "Felix Gabay"
format: html
---

# Problem Statement

Players in the popular MMORPG Old School Runescape often flip commodities on the
game's auction house (the Grand Exchange, or 'GE') for profit.

The age-old advice for newcomers to the flipping market is short and simple.

Buy low, sell high. 

Current strategies for flipping on the GE are mostly heuristic, relying on 
experimentation and data collection until a player finds their rhythm and 
preferred commodities. I have personally explored flipping to some degree, yet
never found enough interest (nor profit) to continue my endeavors. I can't help
but wonder, **does there exist a theoretical best portfolio in terms of profit
margin and sale volume?**


## Sierra 117

Let's start by loading up some packages and the dataset I'll be using.

```{r}
# Packages
packages <- ("tidyverse")

for(pkg in packages) {
  if(!requireNamespace(pkg, quietly = TRUE)) install.packages(pkg)
  library(pkg, character.only = TRUE)
}
```

```{r}
# Data
GE_snapshot_2026_06_27 <- read_csv("../data/GE_snapshot_2026-06-27.csv")
head(GE_snapshot_2026_06_27)
```



